---
title: "p8105_hw3_hw2850"
author: "Haoyang Wang"
date: "2022-10-08"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##problem 2
```{r}
library(tidyverse)
library(dplyr)
accel = read_csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate_if(is.double, as.numeric)
```

```{r}
accel_tidy = 
  pivot_longer(
    accel,
    activity_1:activity_1440,
    names_to = "activity",
    values_to = "activity_data",
    names_prefix = "activity_"
  ) %>% 
  mutate(.data = ., 
         week_type = if_else(day %in% c("Saturday", "Sunday"), "weekend", "weekday"),
         day = fct_relevel(day, "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")) 
```
The tidied dataset has variables `r ls(accel_tidy)`.
The dataset has a size of `r dim(accel_tidy)`.

```{r}
accel_total = accel_tidy %>%  
  group_by(day, week) %>% 
  summarize(activity_day = sum(activity_data)) %>% 
  mutate(week = as.character(week))

accel_table = accel_total %>% table
```

```{r}
ggplot(accel_total, aes(x = day, y = activity_day, group = week, color = week))+
  geom_line()+ 
  scale_y_continuous(
    breaks = c(100000, 400000, 700000),
    labels = c("100000", "400000", "700000"))+
  theme_classic()+
  labs(
    title = "Accelerometer Each Day Data in Five Weeks",
    x = "Day",
    y = "Accelerometer 24-hour Activity",
  )
```
From the graph trend, Accelerometer 24 hour activity is relatively stable in Tuesday, Wednesday.

```{r}
ggplot(accel_tidy, aes(x = activity, y = activity_data, group = day, color = day))+
  geom_line()+
  theme_classic()+
  labs(
    title = "24 Hour Accelerometer Activity for Each Day",
    x = "Activity Count",
    y = "Activity Value",
  )
```
The Accelerator activity of each day fluctuates across a day. 

##problem3
```{r}
library(p8105.datasets)
library(dplyr)
library(rnoaa)
data("ny_noaa")

ny_noaa_tidy = ny_noaa %>% 
  janitor::clean_names() %>% 
  mutate(
    tmax = as.numeric(tmax),
    tmin = as.numeric(tmin),
    tmax = tmax/10,
    tmin = tmin/10
  ) %>% 
  separate(date, into = c("year","month","day"))

ny_noaa_tidy %>% 
  count(snow)
```
The most commonly observed value for snowfall is 0.

```{r}
ny_max = filter(
  ny_noaa_tidy, month %in% c("01", "07")) %>% 
  mutate(
    month = recode(month, `01` = "January", `07` = "July")
  ) %>% 
  select(.data = ., id, year, month, tmax) %>% 
  group_by(id, year, month) %>% 
  summarize(tmax_mean = mean(tmax, na.rm = T)) 
 

ggplot(ny_max, aes(x = year, y = tmax_mean, group = id))+
  geom_line(aes(color = year))+
  scale_x_discrete(
    breaks = c(1980, 1990, 2000, 2010),
    labels = c("1980", "1990", "2000", "2010")
  )+
  theme_classic()+
  facet_grid(. ~ month)+
  labs(
    title = "Max Temperature in January and in July in each NYC Subway Station Across Years",
    x = "Year",
    y = "Average Maximum Temperature (C)"
  )
```
The average maximum temperature fluctuates in January between -10 and 10 degree Celsius and fluctuates in July between 20 and 40 degree Celsius. There are few outliers which could denote extreme weather.

```{r}
ggplot(ny_noaa_tidy, aes(x = tmin, y = tmax), group = week)+
  geom_point(aes(color = year))
```


```{r}
library(ggridges)
ny_snow = ny_noaa_tidy %>% 
  group_by(id, year, snow) %>% 
  summarize(snow = sum(snow)) %>% 
  filter(snow > 0 & snow<100)

ggplot(ny_snow, aes(x = snow, y = year))+
  geom_density_ridges(scale = .85)+
  labs(
    title = "Snowfall From 0 to 100 By Years",
    x = "Snowfall Value",
    y = "Years",
  )
```

